---
title: "Prelim_Spring2019"
author: "Sarah Glancy"
date: "2/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
when importing data set, change mmLenght, Total_prey and Total_gut to numeric values
 how to change row names after import
 
```{r}
library(readxl)
Spring2019 <- read_excel("C:/Users/Sarah/Desktop/WHOI/LTER/201902/LTER201902_FishGuts.xlsx", 
    col_types = c("numeric", "numeric", "text", 
        "numeric", "text", "text", "text", 
        "text", "text", "text", "numeric", 
        "text", "numeric", "numeric", "text", 
        "text"))
View(Spring2019)
```
 
 Spring2019 %>% 
  rename(
    Length(mm) = mmLength )

```{r}


kable(Spring2019)
```

I need to seperate this table into two, one that has unique identifiers for prey and another that had fish with counts from each prey type it contains. 

selecting out specific rows and then naming a new data file with just those rows in them. 
Example: portland_flights <- flights %>% 
            filter(dest == "PDX")
            View(portland_flights)
```{r}
FishCount <- Spring2019 %>% 
  filter(Total_prey != "NA") %>%
           filter(GutItem != "Parasite")
View(FishCount)

```

create another data table with only the rows that have unique prey ID with measurements 
```{r}

PreyMeas <- Spring2019 %>% 
  filter(Prey_mm != "NA")
View(PreyMeas)
```

finally create table that has parasite info- will be remainder of rows. 

```{r}
Parasites <- Spring2019 %>% 
  filter(GutItem == "Parasite")
View(Parasites)

```

