---
title: "Prelim_Spring2019"
author: "Sarah Glancy"
date: "2/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
when importing data set, change mmLenght, Total_prey and Total_gut to numeric values
 how to change row names after import
 
```{r}
library(readxl)
Spring2019 <- read_excel("C:/Users/Sarah/Desktop/WHOI/LTER/201902/LTER201902_FishGuts.xlsx", 
    col_types = c("numeric", "numeric", "text", 
        "numeric", "text", "text", "text", 
        "text", "text", "text", "numeric", 
        "text", "numeric", "numeric", "text", 
        "text"))
View(Spring2019)
```
 
 
 Warnings in above are from the NAs in the numeric columns
 
 Spring2019 %>% 
  rename(
    Length(mm) = mmLength )

```{r}


kable(Spring2019)
```

I need to seperate this table into two, one that has unique identifiers for prey and another that had fish with counts from each prey type it contains. 

selecting out specific rows and then naming a new data file with just those rows in them. 
Example: portland_flights <- flights %>% 
            filter(dest == "PDX")
            View(portland_flights)
```{r}
FishCount <- Spring2019 %>% 
  filter(Total_prey != "NA") %>%
           filter(GutItem != "Parasite")
View(FishCount)

```

create another data table with only the rows that have unique prey ID with measurements 
```{r}

PreyMeas <- Spring2019 %>% 
  filter(Prey_mm != "NA")
View(PreyMeas)
```

finally create table that has parasite info- will be remainder of rows. 

```{r}
Parasites <- Spring2019 %>% 
  filter(GutItem == "Parasite")
View(Parasites)

```

Starting some analysis

n() counts the number of ROWS in each each group

```{r}

by_FSpecies <- Spring2019 %>% 
  group_by(FishSpecies) %>% 
  summarize(count = n())
by_FSpecies
```

###
start attempting data summarizing

```{r}

summary_FishSp_TotGut <- Spring2019 %>% 
  group_by(FishSpecies) %>% 
  summarize(mean = mean(Total_gut, na.rm = TRUE))
summary_FishSp_TotGut
```
 there is a misspelling where one is capatilized in P_tri attempting to change to all match
 
```{r}
install.packages(mudata2)

recode(Spring2019$FishSpecies, p_tri = "P_tri" )

summary_FishSp_TotGut <- Spring2019 %>% 
  group_by(FishSpecies) %>% 
  summarize(mean = mean(Total_gut, na.rm = TRUE))

summary_FishSp_TotGut
```
 rename the "mean" column
 
```{r}

rename (summary_FishSp_TotGut, "Mean prey items" = "mean")

summary_FishSp_TotGut
```
```{r}
print(summary_FishSp_TotGut)
```
more specific groupings
```{r}
summary_FishSp_preytype <- PreyMeas %>% 
  group_by(GutItem) %>%
  summarize(mean = mean(Prey_mm, na.rm = TRUE))

 summary_GutItem <- summary_FishSp_preytype
 
 summary_GutItem
```
requires renaming or "recoding" some of the variables again. 

```{r}

recode(FishCount$GutItem, candacia = "Candacia", polychaeta ="polychaeta" )

summary_GutItem

```
count the number of fish for each species. might need to create a uniqie identifier for each fish??

data(brussels_reviews_anno)
x <- brussels_reviews_anno
x$doc_sent_id <- unique_identifier(x, fields = c("doc_id", "sentence_id"))
head(x, 15)
range(x$doc_sent_id)
x$doc_sent_id <- unique_identifier(x, fields = c("doc_id", "sentence_id"), start_from = 10)
head(x, 15)
range(x$doc_sent_id)

```{r}
install.packages("udpipe")
Spring2019$FishUniq <- unique_identifier(Spring2019, fields = c("Cruise", "Station", "FishSpecies", "FishNum"))

head(Spring2019, 15)

range(Spring2019$FishUniq)

 Spring2019 %>%
   count(Cruise, Station, FishSpecies, FishNum )

```
created a column of unique identifiers, but the total is not equal to the number of fish sampled. 
```{r}

by_FSpecies <- Spring2019 %>% 
  group_by(FishUniq) %>% 
  summarize(count = n())
by_FSpecies

summary(Spring2019$FishUniq)
```
making a smaller tibble and then joining later using my new unique identifier

```{r}

NoPrey_table <-  Spring2019 %>% 
  distinct(FishSpecies, Station, FishNum, .keep_all = TRUE) %>%
  count(FishSpecies)

Spring2019 %>%
  count(FishSpecies, Station, FishNum)
  
NoPrey_table
```

there are still dumplicates, i'll try to find out of it is the station that is somehow not unique
```{r}
recode(FishCount$FishSpecies, p_tri = "P_tri" )

##recode(Spring2019$FishSpecies, p_tri = "P_tri" ) ## this worked above, why not here

Summary_FishCount <- Spring2019 %>% 
  group_by(FishSpecies) %>%
  summarize(count = n())

 Summary_FishCount
```
group by and then mutate to add variable to use for statistics
split to create a new dataframe
long to wide transformation?
```{r}
Spring2019_wide <- Spring2019 %>%
  pivot_wider(
  id_cols = FishSpecies,
  names_from = FishUniq,
  names_prefix = "",
  names_sep = "_",
  names_glue = NULL,
  names_sort = FALSE,
  names_repair = "check_unique",
  values_from = Total_gut,
  values_fill = NULL,
  values_fn = NULL
)

Spring2019 %>%
  pivot_wider(id_cols = FishSpecies, 
              names_from = FishUniq,
              values_from = Total_gut
              
              )

```

```{r}
Spring2019 %>%
  select(FishUniq, FishSpecies, Station) %>%
  group_by(FishUniq, FishSpecies) %>%
  count()


  
```
###Notes from Bret Longworth, 02-19-2021
#green is his original, and active code is my adaptatons

---
title: "Fish Counts"
output:
  html_document:
    df_print: paged
---

Load libraries and data

```{r}
#library(tidyverse)
#library(readxl)
#library(gt)

#df <- read_excel("~/Desktop/LTER201902_FishGuts.xlsx")
#str(df)
#df 
str(Spring2019)
#df 

```

# How many fish of each species?

Select relevant fields. Each of these fish has duplicate rows for each gut item. For counting, we need one row per fish. Finally, if we count the number of each species, that should give us the number by species.

```{r}
#df %>%
  #select(Station, FishSpecies, FishNum) %>% 
  #distinct() %>% 
  #count(FishSpecies)

Spring2019 %>%
  select(Station, FishSpecies, FishNum) %>% 
  distinct() %>% 
  count(FishSpecies)



```
create a bar plot of the fish per species. ## this is still not working

```{r}

#FishSpDist <- Spring2019 %>%
  #select(Station, FishSpecies, FishNum) %>% 
  #distinct() %>% 
  #count(FishSpecies) 

#ggplot(FishSpDist, aes(FishSpecies), ymax = 35) +
  #geom_bar() 
  
```


# How many fish per station?

```{r}
#df %>%
  #select(Station, FishSpecies, FishNum) %>% 
  #distinct() %>% 
  #count(Station)

Spring2019 %>%
  select(Station, FishSpecies, FishNum) %>% 
  distinct() %>% 
  count(Station)


```

# How many fish of each species at each station

I would like to make a stacked bar plot, with either the stations across the bottom or the fish species, so we can see the distribution of species between stations

```{r}
#df %>%
  #select(Station, FishSpecies, FishNum) %>% 
  #distinct() %>% 
  #count(Station, FishSpecies)

Spring2019 %>%
  select(Station, FishSpecies, FishNum) %>% 
  distinct() %>% 
  count( FishSpecies, Station)
```

# How many fish of each species had eaten each prey type?

No longer need `distinct()` to get individual fish because we're now interested in gut content. We can still use `distinct()` to give unique prey types instead of "number of samples that are GutItem.

##the below needs to be further refined, it's too much. -SG

```{r}
#df %>%
  #select(Station, FishSpecies, FishNum, GutItem) %>% 
  #count(FishSpecies, GutItem)

Spring2019 %>%
  select(Station, FishSpecies, FishNum, GutItem) %>% 
  count(FishSpecies, GutItem)
```

```{r}
#df %>%
  #select(Station, FishSpecies, FishNum, GutItem) %>% 
  #distinct() %>% 
  #count(FishSpecies, GutItem)

Spring2019 %>%
  select(Station, FishSpecies, FishNum, GutItem) %>% 
  distinct() %>% 
  count(FishSpecies, GutItem)
```

To make this a more human-readable table, we can pivot to a wider table, with columns for each FishSpecies. This also tells us that the GutItem column needs some recoding.

```{r}
#df %>%
  #select(Station, FishSpecies, FishNum, GutItem) %>% 
  #distinct() %>% 
  #count(FishSpecies, GutItem) %>% 
  #pivot_wider(names_from = FishSpecies, values_from = n) %>% 
  #arrange(GutItem)

recode(Spring2019$GutItem, candacia = "Candacia", CHaetognatha = "Chaetognatha")

Spring2019 %>%
  select(Station, FishSpecies, FishNum, GutItem) %>% 
  distinct() %>% 
  count(FishSpecies, GutItem) %>% 
  pivot_wider(names_from = FishSpecies, values_from = n) %>% 
  arrange(GutItem)
```

##Using recode() on the gut item column to fix the duplicates. should also look at station to make sure there are no duplicates there. -SG

```{r}

 

```


To start on the stats questions, here's the above table broken out by station. I'm using the GT library for the table, which is great for publication-quality tables!


```{r}
#df %>%
  #select(Station, FishSpecies, FishNum, GutItem) %>% 
  #distinct() %>% 
  #group_by(Station) %>% 
  #count(FishSpecies, GutItem) %>% 
  #pivot_wider(names_from = FishSpecies, values_from = n) %>% 
  #arrange(Station, GutItem) %>% 
  #gt()



library(gt)

?gt()


Spring2019 %>%
  select(Station, FishSpecies, FishNum, GutItem) %>% 
  distinct() %>% 
  group_by(Station) %>% 
  count(FishSpecies, GutItem) %>% 
  pivot_wider(names_from = FishSpecies, values_from = n) %>% 
  arrange(Station, GutItem) %>% 
  gt()

```
### End Bret's notes
